---
title: "Stuctural Bioinformatics Class 11"
author: Diego Quintero
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PDB database composition statistics

Download PDB stats as a CSV file from: http://www.rcsb.org/stats/summary

```{r}
pdbstats <- read.csv("Data Export Summary.csv", row.names =1)
```
Let's look at table

```{r}
library(knitr)
kable(pdbstats)
```
Total number of structures 

```{r}
#total number of entries
nstru <- sum(pdbstats[,5])

#percentage of X-Ray structures
perXray <- pdbstats["X-Ray", "Total"]/nstru
perXray <- round(perXray*100,2)

#percentage of Electron Microscopy
perElecMic <- pdbstats["Electron Microscopy","Total"]/nstru
perElecMic <- round(perElecMic*100,2)
```

A1. There are `r perXray`% X-ray structures and `r perElecMic `% Electorn Microsopy structures in the PDB database as of `r Sys.time() `


Calculating the percentage of proteins
```{r}
numProt <- sum(pdbstats$Proteins)/nstru

numProt <- round(numProt*100,2)

```
Total percent of Proteins is `r numProt `%

## Introduction to Bio3D

Loading #Bio3d# library
```{r}
library(bio3d)
```

```{r}
pdb <- read.pdb("1hsg")
pdb
```
Note that the attribute (+attr) of this object are listed on the last couple of lines. To find the attributes of any such object you can use:

```{r}
attributes(pdb)
```

We can access each comonent with the dollar syntac. For Example:
```{r}
head(pdb)
```

```{r}
pdb$xyz
```

Plot of B-factor column... **GOT AN ERROR:( **
```{r}
#plot.bio3d(pdb$atom$b[pdb$calpha], sse=pdb, typ="l", ylab=“B-factor”)
```


##Atom Selection

We need to be able to get at select parts ofprotein structures (e.g just the ligand, just the protein, or just the chain A, etc.). For this we use the **atom.select()** function.


```{r}
ca.inds <- atom.select(pdb, "calpha")

head(pdb$atom[ca.inds$atom,])
```



```{r}
#Download some example PDB files
ids = c("1TND_B", "1AGR_A", "1TAG_A", "1GG2_A","1KJY_A", "4G5Q_A")

files <- get.pdb(ids, split = TRUE)
```

```{r}
# Extract and align the chains we are interested in
pdbs <- pdbaln(files, fit = TRUE)
# Print to screen a summary of the 'pdbs' object
pdbs
```
```{r}
# Access the first 5 rows, and 8 columns
pdbs$ali[1:5, 1:8]


# Associated residues numbers
pdbs$resno[1:5, 1:8]

```

```{r}
# find invariant core
core <- core.find(pdbs)
# superimpose all structures to core
pdbs$xyz = pdbfit(pdbs, core)
# Perform PCA
pc.xray <- pca(pdbs)

# Calculate RMSD
rd <- rmsd(pdbs)
# Structure-based clustering
hc.rd <- hclust(dist(rd))
grps.rd <- cutree(hc.rd, k=3)
plot(pc.xray, 1:2, col="grey50", bg=grps.rd, pch=21, cex=1)

```

