---
title: "Class 08"
author: Diego Quintero
date: 10/25/18
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## K-means clustering

Our first example with **kmeans()** function.

```{r}
#Example plot to see how Rmarkdown works
plot(1:10, typ="l")
```
Back to kmeans...

```{r}
# Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)

k <- kmeans(x, centers=2, nstart=20)

```
Use the kmeans() function setting k to 2 and nstart=20
 Inspect/print the results
 Q. How many points are in each cluster?
```{r}
print(k$size)
```
 
 Q. What ‘component’ of your result object details
   - cluster size?
   `k$size`
   - cluster assignment/membership?
   `k$cluster`
   - cluster center?
  
```{r}
table(k$cluster)
```

```{r}
k$centers
```

   
 Plot x colored by the kmeans cluster assignment and
 add cluster centers as blue points
 
```{r}
palette(c("blue", "red"))
plot(x, col=k$cluster)
points(k$centers, col="blue", pch=20, cex=3)

```
 
 Q. Repeat for k=3, which has the lower tot.withinss? 
 
```{r}
k2 <- kmeans(x, centers = 2, nstart= 20)
k3 <- kmeans(x, centers = 3, nstart= 20)
k4 <- kmeans(x, centers = 4, nstart= 20)
k5 <- kmeans(x, centers = 5, nstart= 20)
k2$tot.withinss
k3$tot.withinss
k4$tot.withinss
k5$tot.withinss

plot(c(k2$tot.withinss, k3$tot.withinss, k4$tot.withinss, k5$tot.withinss))
```
 
# Hierarchical clustering in R

Let's try out the **hclust()** funciton for Hierarchical clustering in R. This function needs a distance matrix as input!


```{r}
d <- dist(x)
hc <- hclust(d)
plot(hc)
#Drae a line on the dendogram
abline(h=6, col="red")

#Cute the tree to yield membership vector at height 
cutree(hc, h=6)

#Cut the tree to yield membership vector by number of groups
cutree(hc, k=2)
```


###A more 'real' example of data clustering

```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
# Step 2. Plot the data without clustering
plot(x)
# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )

palette(c("green", "blue","red"))
plot(x, col=col, pch=20)
```

# Q. Use the dist(), hclust(), plot() and cutree()
# functions to return 2 and 3 clusters

```{r}
dist_matrix <- dist(x)

hc2 <- hclust(dist_matrix)

#plot(hc2)

#Cutting tree into 2 clusters
by_2 <- cutree(hc2, k=2)


#Cutting tree into 3 clusters
by_3 <- cutree(hc2, k=3)

plot(x,col=by_3, pch=20)
```


# Q. How does this compare to your known 'col' groups?

```{r}
#Use the table() function
table(by_3, col)
```

#PCA, Principal Component analysis

Let's make up some data with quite a few dimensions


```{r}
## Initialize a blank 100 row by 10 column matrix
mydata <- matrix(nrow=100, ncol=10)
## Lets label the rows gene1, gene2 etc. to gene100
rownames(mydata) <- paste("gene", 1:100, sep="")
## Lets label the first 5 columns wt1, wt2, wt3, wt4 and wt5
## and the last 5 ko1, ko2 etc. to ko5 (for "knock-out")
colnames(mydata) <- c( paste("wt", 1:5, sep=""),
 paste("ko", 1:5, sep="") )
## Fill in some fake read counts
for(i in 1:nrow(mydata)) {
 wt.values <- rpois(5, lambda=sample(x=10:1000, size=1))
 ko.values <- rpois(5, lambda=sample(x=10:1000, size=1))

 mydata[i,] <- c(wt.values, ko.values)
}
head(mydata)
```


Note the prcomp() 
```{r}
head(t(mydata))
```

Now let's try to find "structure" in this data

```{r}
pca <- prcomp(t(mydata), scale=TRUE)
pca

#a basic PC1 vs PC2 2-D plot
plot(pca$x[,1],pca$x[,2])

## Variance captured per PC
pca.var <- pca$sdev^2 

## Precent variance is often more informative to look at
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)

pca.var.per
```


Make a "scree-plot" to see the variance captured in each PC

```{r}
barplot(pca.var.per, main="Scree Plot",
 xlab="Principal Component", ylab="Percent Variation")
```

```{r}
colvec <- colnames(mydata) 
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"

plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
 xlab=paste0("PC1 (", pca.var.per[1], "%)"),
 ylab=paste0("PC2 (", pca.var.per[2], "%)")) 


## Lets focus on PC1 as it accounts for > 90% of variance
loading_scores <- pca$rotation[,1] 
summary(loading_scores)


## We are interested in the magnitudes of both plus
## and minus contributing genes
gene_scores <- abs(loading_scores) 

```
#Reading UK Food datausing PCA

```{r}
x <- read.csv("UK_foods.csv")
dim(x)

head(x)

#We were expecting 4 columns, rather than set as proper row names
#Fix this with code below
rownames(x) <- x[,1]
x <- x[,-1]
head(x)

dim(x)

knitr::kable(x, caption="The full UK foods data table")
```





